---
title: "Probability & Statistic Final Assignment"
author: "Tuan Hoang^[IS1412, tuanhmhe141540@fpt.edu.vn]"
date: "03/18/2022"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

## Setup

### Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(gssr)
```

### Load data

```{r load-data, message=FALSE}
data(gss_doc)
gss21 <- gss_get_yr(2021)
```

------------------------------------------------------------------------

## Section 1: Data

### Overview

The General Survey (also known as GSS) has been conducted since 1972 to
cater to statisticians and data scientists discovering more about public
opinion trends in the US.

### Sampling Design

The collecting data method in 2021 was implemented significant
methodological changes to ensure the safety of both respondents and
interview. For instances, moving from in-person to address-based
sampling and online methodology.

Let's go through the details

In order to conduct the GSS, agents obtain samples of web-based and
address-based survey <br> <br>

<ul>

<li>

<b>The mail-based Sample</b><br>

<p>

Used DSS - Disproportionate stratified sampling since 2006. They got
addresses from strata that were based on NORC's master sample and their
charateristics such as: the the density of the household which is from
the America Community Survey. Samples are taken from the north-west
corner inward until quotas are met.

In that design, addresses were classified into 2 kinds of stratum that
either was metropolitan or rura and within it new primary sampling units
(PSUs) are being used to yield addresses for surveying.

</p>

</li>

<li>

<b>The Web-based Sample</b><br>

<p>

As an online questionnaire. In this survey, some of the questions that
had the same theme were listed on the screen together as a grid so that
they can answer them in one go.

Before being published, the data were weighted based on the proportions
to equal the 2019 American Community Survey Census Bureau to ensure
unbiased characteristics.

</p>

</li>

</ul>

### Scope of inference

As mentioned above, GSS use stratified sampling design based on the NORC
master sample to select a representative sample from each state
residents.

Thus, each state began with a single stratum to provide adequate sample
sizes which is defined populations of interest but many states sample
were disproportionate in their strata that correspond to sub-state or
non-metropolitant regions.

### Generalizability

The data are collected from multiple areas. Moreover, in each cluster
group is sampled, the measure is taken at higher levels and by
percentages to ensure the distribution of key variables (sex, gender,
etc).

Therefore, it can be generalized to the US population.

### Causality

Making causal conclusions based on observational data is not
recommended. Observational studies are only sufficient to show
correlations.

### Summary

The data collection is based on a stratified sampling strategy. The
population is divided into groups (Region) called strata. Then a random
sampling (characteristics), has been employed within each stratum.It can
be generalized to the US population

<b>The data and further information were used in this project from the
following sources:</b>

General Information: [2021 Survey Data
Information](http://www.cdc.gov/brfss/annual_data/annual_2013.html)

------------------------------------------------------------------------

## Section 2: Research questions

We would like to focus on the behaviors, demographics, and opinions

**Research question 1:**

Do women oftentimes feel uneasy strolling alone at night in their
neighborhood?

Most people often hold the common perception that the girls are more
afraid of walking alone at night than the males do.

We can further dive into it by splitting the data by geographical
locations

-   Explanatory variable: `sex`

```{r}
  gss_doc %>% filter(id == "sex") %>% 
  select(description)
```

-   Response variable: `fear`

```{r}
  gss_doc %>% filter(id == "fear") %>% 
  select(description)
```

**Research question 2:**

**Research question 3:**

What is the relationship between the age of surveyed person and the
hours he/she watches a day in average?

-   Explanatory variable: `age`

```{r}
  gss_doc %>% filter(id == "age") %>% 
  select(description)
```

-   Response variable: `tvhours`

```{r}
  gss_doc %>% filter(id == "tvhours") %>% 
  select(description)
```

------------------------------------------------------------------------

## Section 3: EDA - Exploratory data analysis

**Research question 1:** ***Do women oftentimes feel uneasy strolling
alone at night in their neighborhood?***

First, we need to get the data and clean to get the grasp of the data

```{r load-ffear-data}
gss21_sex <- 
  gss21 %>% filter(!is.na(gss21$sex), is.numeric(gss21$sex)) %>% 
  mutate_at(vars(matches("sex")) , ~ifelse(sex == 1 , "Male", "Female")) 

ffear <- 
  gss21_sex %>% filter(sex == "Female", !is.na(gss21_sex$fear), is.numeric(gss21_sex$fear)) %>%
  mutate_at(vars(matches("fear")), ~ifelse(fear == 1, "Yes", "No"))

ffear %>% group_by(sex, fear) %>% summarise(count = n())
```

What I did there is filter out the "NA" entries and convert fields to
human-readable format.

As depicted from the table, there are a significant amount of
observations that women are comfortable strolling at night. This also
indicates that most girls are all right to take a walk in dark.

Let's further investigate the statement

```{r plot-ffear-sample}
ffear$fear <- as.factor(ffear$fear)
ggplot(ffear, aes(x = fear)) +
  geom_bar(fill = "#8FDEE1") +
  xlab("Fear") +
  ylab("Number of people") + 
  ggtitle('Sample Distribution')
```

As is evident from the graph, the number of girls who are not afraid
when walking at night outweighed the scared ones.

**Research question 3:** ***What is the relationship between the age of
surveyed person and the hours he/she watches a day in average?***

First, we need to get the data and clean to get the grasp of the data

```{r load-ftvhours-data}
gss21_age <- 
  gss21 %>% filter(!is.na(gss21$age), is.numeric(gss21$age)) %>% 
  mutate_at(vars(matches("age")) , ~ifelse(is.na(age) , -1, age))

ftvhours <- 
  gss21_age %>% filter(!is.na(gss21_age$tvhours), is.numeric(gss21_age$tvhours)) %>%
  mutate_at(vars(matches("tvhours")) , ~ifelse(is.na(tvhours) , -1, tvhours))

ftvhours %>% group_by(age, tvhours) %>% summarise(count = n())
```

```{r plot-ftvhours-sample}
ggplot(ftvhours, aes(x = age, y = tvhours)) +
  geom_point(fill = "#8FDEE1") +
  xlab("Age") +
  ylab("tv hours") + 
  ggtitle('Sample Distribution')
```

------------------------------------------------------------------------

## Section 4: Inference

**Research question 1:**

***Confidence Interval:***

*Conditions for central limit theorem:*

1.  Independence Check: The observations are collected from a simple
    random sample. Also, it consists of fewer observations than 10% of
    the US population, which we could safely assume to be independent.
2.  Sample size / skew check (success-failure check): The sample size
    `n` of our dataset was 2,197. From which 1,239 success (not-afraid)
    cases and 958 failure (scare) cases.

Both conditions passed, thus the distribution is nearly normal as
defined by the central limit theorem.

We will construct a 95% CI for the proportion of girls that are not
afraid to walk alone in the evening in their locals

```{r compute-ffear-ci}
ffear_len <- length(ffear$fear)
ffear_p_hat <- sum(ffear$fear == "No") / ffear_len

ffear_conf_level <- 0.95
ffear_perc_crit_value <- ffear_conf_level + ((1 - ffear_conf_level) / 2)

ffear_zstar <- qnorm(ffear_perc_crit_value)

ffear_se <- sqrt(ffear_p_hat * (1 - ffear_p_hat) / ffear_len)
ffear_me <- ffear_zstar * ffear_se

ffear_ci  <- ffear_p_hat + c(-1,1) * ffear_me


cat(paste0("n = ", ffear_len, ", p-hat = ", round(ffear_p_hat, 4), "\n",
ffear_conf_level, "% CI: (", round(ffear_ci[1], 4), " , ", round(ffear_ci[2], 4), ")\n"))

```

We are `r ffear_conf_level * 100`% confident that
`r round(ffear_ci[1], 4)`% to `r  round(ffear_ci[2], 4)`% of all
American females are ok with night strolls.

***Hypothesis testing:***

Let's continue the question from above: We have 2 hypothesizes.

*H0:* Girls are terrified of walking without accompany at night in their
near vicinity (p = 0.5)\
*H1:* They walk at night in their neighborhood without hesitant (p \>
0.5)

*Conditions for central limit theorem:*

1.  Independence Check: As mentioned above, we can safely assume it
    independence

2.  Sample size / skew check (success-failure check): From the data, we
    see that the observed success rate (p-hat) is `r ffear_p_hat` <br/>
    The test proportion `p` used in our hypotheses will be 0.5 <br/> We
    also had np = `r ffear_len` \* 0.5 = `r ffear_len * 0.5`

Those gave us more than 10 cases for both successes and failures

Because all the tests passed, the distribution is nearly normal as
defined by the central limit theorem.

Let's look into this.

```{r compute-ffear-ht}
ffear_null = 0.5
ffear_zstats <- (ffear_p_hat - ffear_null) / ffear_se

ffear_x_min <- ffear_p_hat 
ffear_x_max <- Inf

ffear_p_value <- pnorm(ffear_zstats, lower.tail = FALSE)
ffear_p_val_to_print <- ifelse(round(ffear_p_value, 4) == 0, "< 0.0001", round(ffear_p_value, 4))

cat(paste0("n = ", ffear_len, ", p-hat = ", round(ffear_p_hat, 4), "\n",
           "H0: p = ", ffear_null, "\n" ,
          "HA: p >", " ", ffear_null, "\n",
          "z = ", round(ffear_zstats, 4), "\n",
          "p_value = ", ffear_p_val_to_print
          ))
```

The hypothesis test yields `r ffear_zstats`, making the p-value
negligible (`r ffear_p_val_to_print`).

Let's construct the null sample graph to make it more transparent

```{r plot-ffear-null}
ggplot(data.frame(x = c(ffear_null - 4 * ffear_se, ffear_null + 4 * ffear_se)), aes_string(x= 'x')) + 
  stat_function(fun = dnorm, args = list(mean = ffear_null, sd = ffear_se), color = "#999999") +
  annotate("rect", xmin = ffear_x_min, xmax= ffear_x_max, ymin = 0, ymax = Inf, alpha = .3, fill= "#FABAB8") + 
  ggtitle("Null distribution") +
  xlab("") +
  ylab("") +
  geom_vline(xintercept = ffear_p_hat, color ="#F57670", lwd = 1.7)
```

By observing the results, we can see that the probability of the null
hypothesis is almost 0. Therefore, we reject the null hypothesis.

------------------------------------------------------------------------

## Section 5: Prediction

**Research question 3:** ***What is the relationship between the age of
surveyed person and the hours he/she watches a day in average?***

```{r linear-regression}
lmpredict = lm(ftvhours$tvhours~ftvhours$age)
summary(lmpredict)
```

```{r predict-plot}

predict_df <- data.frame(tvhours_pred = predict(lmpredict, ftvhours), age=ftvhours$age)

ggplot(ftvhours, aes(x = age, y = tvhours)) +
  geom_point(fill = "#8FDEE1") +
  geom_smooth(method = lm, color="red")
  xlab("Age") +
  ylab("tv hours") + 
  ggtitle('Sample Distribution')
```

------------------------------------------------------------------------

## Section 6: Conlusion

**Research question 1:**

From our hypothesis testing results, we can accept the alternative
hypothesis. That is, that females are afraid of walking alone at night
in their local neighborhood. As mentioned previously, further research
could be done to analyze this response by geographical location to see
if this result varies across different regions in America.

**Research question 3:**

By using Linear Regression, we can get the relationship between *the age
of surveyed person and the hours he/she watches a day in average is a
line by [tvhours] = 0.035377 \* [age] + 1.631708*

------------------------------------------------------------------------

## Section 7: Citations & References

Healy K (2019). gssr: General Social Survey data for use in R. R package
version 0.2.0, <http://kjhealy.github.io/gssr>.

GSS website: <http://gss.norc.org/> (11/01/2021)
